---
title: "Presliced hams"
author: "Marinko Maslaric & Dejan Mircetic"
date: "11/03/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.pos = 'H')
library(readxl)
library(tidyverse)
library(tsibble)
library(lubridate)
library(fable)
library(thief)
```

<!-- ## Structuring the data -->

```{r import}
sliced <- read_excel("1. Draft/Draft Data structure/Data 2 februar 2022/PRESLICED FILE.xls")
```

Data has `r dim(sliced)[1]` entries, from which not all are reliable and need to be checked again.

```{r seting_up, include=FALSE}

sliced <- sliced %>%
  filter(!is.na(DATE))

colnames(sliced) <- c("date","doc","product_ID","product_type","curing","pices","kg","value","pos")

sliced %>% select(date,product_type,pos,curing,pices,kg,value)

base <- sliced  %>% 
  select(date,product_type,pos,curing,pices,kg,value) %>% 
  mutate(date=ymd(date), curing=as.character(curing), kg=as.numeric(kg), value=as.numeric(value)) 

agregiranje <- base %>% 
  duplicates(key = c(pos,product_type, curing), index=date) %>%
  group_by(date, pos,product_type, curing) %>% 
  summarise(kg = sum(kg), pices=sum(pices), value=sum(value))
```


```{r duplicates}
## Checking for duplicates
# Logicki vektori a i b za izbacivanje iz baze dupliranih podataka
a <- base %>% 
  are_duplicated(key = c(pos,product_type, curing), index=date)

b <- base %>% 
  are_duplicated(key = c(pos,product_type, curing), index=date,from_last = TRUE)

#bind_rows(base[!c(a|b),],agregiranje) %>% 
#  duplicates(key = c(pos,product_type, curing), index=date)

#Kalemljenje      
baza <- bind_rows(base[!c(a|b),],agregiranje)
```


```{r}
## Joinig the data & NA values
#Kalemljenje  
baza <- bind_rows(base[!c(a|b),],agregiranje)

baza <- baza %>%
  as_tsibble(key =  c(pos,product_type, curing), index=date) %>%
  fill_gaps(kg = 0L, pices = 0L, value = 0L)

#baza %>%
#      filter(is.na(product_type))
```

Demonstrating the demand for one customer:

```{r fig1}
baza %>%
  filter(pos=="70.101.893 SOGEGROSS" & curing=="18") %>%
  autoplot(value)+
  ggtitle("Fig 1. Visualisation of presliced hams SOGEGROSS customer")
```

## Total demand of presliced hams (all customers)

Fig. 2 represents the total number of presliced hams which were sold/delivered to the market from 2015.
```{r, out.width="100%"}
## Creating the grouped data base
data_agg <- baza %>%
  aggregate_key(pos * product_type* curing, Hams = sum(kg), Pices = sum(pices), Value = sum(value))

total_level <- data_agg %>%
  filter(is_aggregated(pos) & is_aggregated(product_type) & is_aggregated(curing)) #Za tekst

maxx <- which.max(total_level[,"Pices"]$Pices)
datum <- total_level[,"date"]$date[maxx]

data_agg %>%
  filter(is_aggregated(pos) & is_aggregated(product_type) & is_aggregated(curing)) %>%
  autoplot(Pices) +
  xlab("Days")+
  geom_vline(aes(xintercept=as.Date("2020-02-21"), colour="#990000")) +#Corona start! First lock down
  scale_color_hue(labels = c("Lockdown in Italy"))+
  ggtitle("Fig 2. Total consumption")
```


There is a notable change in the hams demand from the start of the Corona pandemic. The first lockdown in Italy was **2020-02-21**, after which demand for presliced hams erupted and reached a historical maximum. The demand was constantly trending, reaching the highest peak on **`r datum`**, with **`r format(max(total_level[,"Pices"]),scientific=FALSE)`** sliced pieces sold and generated revenue of **`r format(max(total_level[,"Value"]),scientific=FALSE)`** EUR in one day.

```{r, eval=FALSE}
library(fpp3)
total_level %>% select(Hams) %>%
  gg_season(labels = "both")


total_level %>% select(Hams) %>%
  model(
    STL(Hams ~ trend(window = 7) +
          season(window = "periodic"),
        robust = TRUE)) %>%
  components() %>%
  autoplot()

```

The daily demand shows highly volatile behavior, therefore aggregation of the demand through different time horizons (weekly, monthly, etc) demonstrate a clear galloping demand trend.


```{r, echo=FALSE}
df <- total_level %>% select(date,Hams, Pices, Value)
df <- as_tibble(df)

#head(df)

df$week <- floor_date(df$date, "week")
df$month <- floor_date(df$date, "month")

sedmica <- df %>%
  group_by(week) %>%
  summarize(Consumption = sum(Hams)) 

#Weekly_demand <- ts(sedmica$Consumption,frequency = 52, start = 2015)
#plot(Weekly_demand)

tsw <- sedmica %>% 
  mutate(month=yearweek(week)) %>%
  as_tsibble(index=week)

tsw %>% autoplot(Consumption)+
  xlab("Weeks")

```

Similar behavior is observed on the monthly and other upper levels as well.

```{r}
mjesec <- df %>%
  group_by(month) %>%
  summarize(Consumption = sum(Hams)) 

Monthly_demand <- ts(mjesec$Consumption,frequency = 12, start = c(2015,1))
#plot(Monthly_demand)
#plot(forecast(Monthly_demand))

plot(tsaggregates(Monthly_demand, align = "end"), main="Demand through time horizonts")
```

Initial estimations for the future forecasts are presented in the next figure and still demonstrate strong consumption of the presliced ham on the market.

```{r}
tsm <- mjesec %>% 
  mutate(month=yearmonth(month)) %>%
  as_tsibble(index=month)

#autoplot(tsm,Consumption)

fc <- tsm %>%
  model(
     demo_pessimistic= ARIMA(Consumption),
    #demo_optimistic=ETS(Consumption)
  ) %>%
  forecast(h = 24)

fc %>%
  autoplot(tsm)+
  ggtitle("Total consumption & forecasts")+
  xlab("Months")

#Kad budes pices prognozirao dobija bolju sliku nego hams
```

